---
title: "p8105_hw1_amh2325"
author: "Amanda Howarth"
date: "9/19/2019"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
``` {r}
library(tidyverse)
```
## Problem 1: 
This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

## Creating a dataframe 
This data frame will include classes in a high school and the floor they are located on within the school 
```{r}

#create a dataframe 
set.seed(8)
hw_df = tibble(
    samp = rnorm(8),
    logic = samp > 0,
    class = c("biology", "chemistry", "physics", "history", "math", "art", "music", "gym"),
    floor = factor(c("first", "second", "third", "second", "first", "second", "third", "first"))
)

```
Taking the mean of each variable
```{r}
#taking mean values for each variable in hw_df 
mean(pull(hw_df, samp))
mean(pull(hw_df, logic))
mean(pull(hw_df, class))
mean(pull(hw_df, floor))

```
We received the mean for the numeric (samp) and logical (logic) variables. For both the charcter (class) and factor (floor) variables, R would not allow us to get the mean. We receive the following warning: argument is not numeric or logical: returning NA[1] NA. According to r, This is because "there are currently methods for numeric/logical vectors and date, date-time, and time interval objects. Complex vectors are allowed for trim=0." A mean cannot be obtained for the character and factor variables since they are technically not numeric.
```{r}
as.numeric(pull(hw_df,logic))
as.numeric(pull(hw_df,class))
as.numeric(pull(hw_df,floor))
```
When we run the above codes, the values are outputted for the "logic" and "floor" variable but not for the "class" variable. This is because the logical variable, "logic," has a value of 0 (false) and 1(true). When the factor variable (categorical ordinal) was converted into numeric, the "floor" variable was assigned numbers according to its order: 1=first, 2=second, 3=third. However, the "class" variable (categorical nominal) would still not produce a mean because it cannot be converted to a numeric variable. This example explains why a mean can only be produced for variables that are numeric or logical. 

Converting logical variables 
```{r}
#converting logical vector to numeric and multiplying by samp
logic = as.numeric(pull(hw_df, logic))
pull(hw_df, samp)*logic 
```

```{r}
#converting logical vector to a factor and then multiplying by samp
logic = as.factor(pull(hw_df, logic))
pull(hw_df, samp)*logic
```
```{r}
#converting logical vector to a factor, then converting the result to numeric, and multiplying by samp
as.factor(pull(hw_df, logic))
as.numeric(as.factor(pull(hw_df, logic)))*pull(hw_df, samp)
```
## Problem 2: 
This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.
```{r}
#creating a dataframe for Problem 2
set.seed(1000)
hw2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logical_vector = c((x + y) > 1),
  numeric_vector = as.numeric(logical_vector),
  factor_vector = as.factor(logical_vector)
  )
```

```{r}
#rows and columns in hw2_df 
  nrow(hw2_df)
  ncol(hw2_df)

#mean, median, and standard deviation of x 
mean(pull(hw2_df, x))
median(pull(hw2_df, x))
sd(pull(hw2_df, x))

#the proportion of cases for which x+y>1 
mean(pull(hw2_df, logical_vector))

#In my vector, there are 500 rows in the dataset and five columns. The mean value of x is 0.008544891, and the median value of x is 0.006517422. The standard deviation of x is 0.9390666. The proportion of cases for which x + y > 1 is 0.236.
```
## Making a scatterplot 
```{r}
#plot 1 with logical_vector 
plot1 <- ggplot(hw2_df, aes(x = x, y = y, color = logical_vector)) + geom_point()

ggsave ("plotdata.pdf", plot = plot1)
```

```{r}
#plot 2 with numeric_vector 
plot2<- ggplot(hw2_df, aes(x = x, y = y, color = numeric_vector)) + geom_point()
```

```{r}
#plot 3 with factor_vector 
plot3 <- ggplot(hw2_df, aes(x = x, y = y, color = factor_vector)) + geom_point()
```
Plot1, which color points using the logical_vector, and Plot3, which color points using the factor_vector, are identical in color scale. The points in the plot are presented in two colors, which represent true (green) and false (pink). This makes sense because the logical_vector is a logical variable (true/false) and the factor_vector variable we created is categorical ordinal, created by coercing the logical variable. PLot 2, which color points using the numeric variable, has a range of blue colors - ranging from 0 (darkest blue) to 1 (lightest blue). This makes sense because the numeric_vector variable was created by coercing logical_vector, and thus, false=0 and true=1. 

