---
title: "p8105_hw1_amh2325"
author: "Amanda Howarth"
date: "9/19/2019"
output: github_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(tidyverse)
```

## Question 1 
## Creating a dataframe 

This data frame will include classes in a high school and the floor they are located on within the school 

```{r}

#create a dataframe 
set.seed(8)
HW_df = tibble(
    samp = rnorm(8),
logic = samp > 0,
class = c("biology", "chemistry", "physics", "history", "math", "art", "music", "gym"),
floor = factor(c("first", "second", "third", "second", "first", "second", "third", "first"))
)

```

Taking the mean of each variable
```{r}
mean(pull(HW_df, samp))
mean(pull(HW_df, logic))
mean(pull(HW_df, class))
mean(pull(HW_df, floor))

```
We received the mean for the numeric (samp) and logical (logic) variables. For both the charcter (class) and factor (floor) variables, R would not allow us to get the mean. We receive the following warning: argument is not numeric or logical: returning NA[1] NA. This is because there are currently methods for numeric/logical vectors and date, date-time, and time interval objects. Complex vectors are allowed for trim=0. 

```{r}
as.numeric(pull(HW_df,logic))
as.numeric(pull(HW_df,class))
as.numeric(pull(HW_df,floor))
```