---
title: "p8105_hw1_amh2325"
author: "Amanda Howarth"
date: "9/19/2019"
output: github_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(tidyverse)
```

## Problem 1: 
This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

## Creating a dataframe 

This data frame will include classes in a high school and the floor they are located on within the school 

```{r}

#create a dataframe 
set.seed(8)
hw_df = tibble(
    samp = rnorm(8),
logic = samp > 0,
class = c("biology", "chemistry", "physics", "history", "math", "art", "music", "gym"),
floor = factor(c("first", "second", "third", "second", "first", "second", "third", "first"))
)

```

Taking the mean of each variable
```{r}
#taking mean values 
mean(pull(hw_df, samp))
mean(pull(hw_df, logic))
mean(pull(hw_df, class))
mean(pull(hw_df, floor))

```
We received the mean for the numeric (samp) and logical (logic) variables. For both the charcter (class) and factor (floor) variables, R would not allow us to get the mean. We receive the following warning: argument is not numeric or logical: returning NA[1] NA. This is because there are currently methods for numeric/logical vectors and date, date-time, and time interval objects. Complex vectors are allowed for trim=0. 

```{r}
as.numeric(pull(hw_df,logic))
as.numeric(pull(hw_df,class))
as.numeric(pull(hw_df,floor))
```
When we run the above codes, the values are outputted for the "logic" and "floor" variable but not for the "class" variable. This is because the logical variable has a value of 0 (false) and 1(true). There factor variable also technically has numeric variables as it is ordered: 1=first, 2=second, 3=third. 

Converting logical variables 
```{r}
#converting logical vector to numeric and multiplying by samp
logic = as.numeric(pull(hw_df, logic))
pull(hw_df, samp)*logic 
```

```{r}
#converting logical vector to a factor and multiplying by samp
logic = as.factor(pull(hw_df, logic))
pull(hw_df, samp)*logic
```
```{r}
#converting logical factor to a factor, then converting the result to numeric, and multiplying by samp

as.factor(pull(hw_df, logic))
as.numeric(as.factor(pull(hw_df, logic)))*pull(hw_df, samp)
```
## Problem 2: 
This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

```{r}

#create a second dataframe for Problem 2
set.seed(1000)
hw2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  logical_vector = c((x + y) > 1),
  numeric_vector = as.numeric(logical_vector),
  factor_vector = as.factor(logical_vector)
  )

nrow(hw2_df)
ncol(hw2_df)
mean(pull(hw2_df, x))
median(pull(hw2_df, x))
sd(pull(hw2_df, x))
mean(pull(hw2_df, logical_vector))
#In my vector, there are 500 rows in the dataset and five columns. The mean value of x is 0.008544891, and the median value of x is 0.006517422. The standard deviation of x is 0.9390666. The proportion of cases for which x + y > 1 is 0.236.

```

## Making a scatterplot 

```{r}
#plot with logical_vector 
plot1 <- ggplot(hw2_df, aes(x = x, y = y, color = logical_vector)) + geom_point()
```

```{r}
#plot with numeric_vector 
plot2<- ggplot(hw2_df, aes(x = x, y = y, color = numeric_vector)) + geom_point()
```

```{r}
#plot with factor_vector 
plot3 <- ggplot(hw2_df, aes(x = x, y = y, color = factor_vector)) + geom_point()
```

Plot1 and Plot2 are identical in color and scatter. PLot 3 is also identical in scatter, however it is has different colors. 


