---
title: "p8105_hw1_amh2325"
author: "Amanda Howarth"
date: "9/19/2019"
output: github_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
library(tidyverse)
```

## Question 1 
## Creating a dataframe 

This data frame will include classes in a high school and the floor they are located on within the school 

```{r}

#create a dataframe 
set.seed(8)
hw_df = tibble(
    samp = rnorm(8),
logic = samp > 0,
class = c("biology", "chemistry", "physics", "history", "math", "art", "music", "gym"),
floor = factor(c("first", "second", "third", "second", "first", "second", "third", "first"))
)

```

Taking the mean of each variable
```{r}
#taking mean values 
mean(pull(hw_df, samp))
mean(pull(hw_df, logic))
mean(pull(hw_df, class))
mean(pull(hw_df, floor))

```
We received the mean for the numeric (samp) and logical (logic) variables. For both the charcter (class) and factor (floor) variables, R would not allow us to get the mean. We receive the following warning: argument is not numeric or logical: returning NA[1] NA. This is because there are currently methods for numeric/logical vectors and date, date-time, and time interval objects. Complex vectors are allowed for trim=0. 

```{r}
as.numeric(pull(hw_df,logic))
as.numeric(pull(hw_df,class))
as.numeric(pull(hw_df,floor))
```
When we run the above codes, the values are outputted for the "logic" and "floor" variable but not for the "class" variable. This is because the logical variable has a value of 0 (false) and 1(true). There factor variable also technically has numeric variables as it is ordered: 1=first, 2=second, 3=third. 

Converting logical variables 
```{r}
#converting logical vector to numeric and multiplying by samp
logic = as.numeric(pull(hw_df, logic))
pull(hw_df, samp)*logic 
```

```{r}
#converting logical vector to a factor and multiplying by samp
logic = as.factor(pull(hw_df, logic))
pull(hw_df, samp)*logic
```
```{r}
#converting logical factor to a factor, then converting the result to numeric, and multiplying by samp

as.factor(pull(hw_df, logic))
as.numeric(as.factor(pull(hw_df, logic)))*pull(hw_df, samp)
```
## Question 2 